---
title: "Musicology_Portfolio"
author: "Ragnar van Dort"
date: "2/22/2022"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: flatly
    
---
```{r setup}
  library(tidyverse)
  library(plotly)
  library(spotifyr)
  library(compmus)
```

### The best music for studying

The main goal of this corpus is to compare playlists, to be specific, compare study playlists. A lot of people study, and a lot of people study while listening to music. All people are different and we also see this in their music taste; it's often said that classical music is great for studying and this is also backed by different studies but there also a lot of other kinds of music that students listen to: LoFi, techno etc.

Over the years I created a study playlist myself, it consist of roughly 3 genres I think: some songs with text but they are not to energetic and are a bit "slow", some techno music and some classical music. I highly doubt it's te the best playlist for studying, but over the years I know that I studied with these songs on. Maybe some songs are not ideal itself to listen to during studying but they help me get in a study mode. Maybe good to know due to it's variety its more a playlist that I use as a storage space for study music than that I actually play the playlist while studying. 


In this corpus I will compare my study playlist with some popular spotify study playlists, I'm curious how different my playlist is compared to a "normal" study playlist. 

I will work with four different playlists all created by spotify: a pop study playlist, a classical music study playlist, a techno study playlist and a LoFi study playlist.

I will compare my own study playlist (stu(dying)) with some popular spotify playlists. For the best comparison I decided to take three spotify playlists with the genres that are also mainly aanwezig in my own playlist:

- Pop Study (sudy music with text): https://open.spotify.com/playlist/37i9dQZF1DWSoyxGghlqv5 
- Classical Studying music: https://open.spotify.com/playlist/37i9dQZF1EIgLoMVUd9oTU 
- Electronic Focus: https://open.spotify.com/playlist/37i9dQZF1DX0wMD4IoQ5aJ 
- My own study music playlist: https://open.spotify.com/playlist/7ex5hFjCiAdaXy6FFoe2Vi

My hypothesis for this research is that there will be a lot of similarities, probably a low bpm and not too energetic for example. But I hope to find some interesting differences that I can then investigate further and possibly explain.

There are a few interesting and atypecal tracks in the playlist. There are of course some tracks that feature in my own playlist and also in spotify's playlist (Una mattine, Comptine d'un autre été: L'après-midi) altough they are a different version. Maybe some atypical music are the movie soundtracks that are in my playlist like: Time-Hans zimmer and the Godfather finale- Nino Rotta.

### The Tempo and Instrumentalness in study music

```{r tempo and intrumentalness}
pop <- get_playlist_audio_features("", "37i9dQZF1DWSoyxGghlqv5")
classic <- get_playlist_audio_features("", "37i9dQZF1EIgLoMVUd9oTU")
electronic <- get_playlist_audio_features("", "37i9dQZF1DX0wMD4IoQ5aJ")
studying <- get_playlist_audio_features("", "7ex5hFjCiAdaXy6FFoe2Vi")

study <-
  pop %>%
  mutate(country = "Pop study") %>%
  bind_rows(classic %>% mutate(country = "Classical study")) %>%
  bind_rows(electronic %>% mutate(country = "Electronic study")) %>%
  bind_rows(studying %>% mutate(country = "Stu(dying)"))%>%
  mutate(
    country = fct_relevel(country, "Pop study, Classical study, Electronical study", "Stu(dying)")
  )

february_dip <-
  study %>%
  ggplot(                          # Set up the plot.
    aes(
      x = tempo,
      y = instrumentalness,
      size = track.popularity,
      colour = energy,
      label = track.name           # Labels will be interactively visible.
    )
  ) +
  geom_point() +                   # Scatter plot.
  geom_rug(size = 0.1) +           # Add 'fringes' to show data distribution.
  facet_wrap(~country) +           # Separate charts per country.
  scale_x_continuous(              # Fine-tune the x axis. Outliers are not 
    limits = c(0, 200),
    breaks = c(0, 100, 200),        # Use grid-lines for quadrants only.
    minor_breaks = c(25, 50, 75, 125, 150, 175)            # Remove 'minor' grid-lines.
  ) +
  scale_y_continuous(              # Fine-tune the y axis in the same way.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c(          # Use the cividis palette
    option = "E",                  # Qualitative set.
    alpha = 0.8,                   # Include some transparency
    guide = "none"
  ) +
  scale_size_continuous(           # Fine-tune the sizes of each point.
    guide = "none"                 # Remove the legend for size.
  ) +
  theme_light() +                  # Use a simpler theme.
  labs(                            # Make the titles nice.
    x = "Tempo",
    y = "Instrumentalness"
  )

ggplotly(february_dip)
```
*** 

There is no consensus in the literature about what music is best for studying but every answer has one thing in common; the main goal is that it should help you to focus. Generally speaking music with a relative low bpm (<120 bpm) is best suited for this. In study music (new) texts also seem to be distractive, so a high instrumentalness would be a good sign. 

I plotted the instrumentalness on the y-axis, the tempo on the x-axis, the color is the energy of the song and the size is its popularity. 

The plots are kind of what I expected. The Spotify music playlists are pretty clustered while my own study playlist is not so much. This is not really hard to explain. The Spotify playlists consist of one genre while my playlist is a mix of genres. 

### Comptine d'un autre été: L'après-midi[chroma features].

```{r tallis}
lavinia <-
  get_tidy_audio_analysis("06MWWxWRuBtYxcJP71tN0q") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
tierssen <-
  get_tidy_audio_analysis("14rZjW3RioG7WesZhYESso") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
maria_dist <-
  compmus_long_distance(
    lavinia %>% mutate(pitches = map(pitches, compmus_normalise, "euclidean")),
    tierssen %>% mutate(pitches = map(pitches, compmus_normalise, "euclidean")),
    feature = pitches,
    method = "cosine"
  )
```

```{r tallis-plot}
maria <-
  maria_dist %>%
  mutate(
    lavinia = xstart + xduration / 2,
    tierssen = ystart + yduration / 2
  ) %>%
  ggplot(
    aes(
      x = lavinia,
      y = tierssen,
      fill = d
    )
  ) +
  geom_tile(aes(width = xduration, height = yduration)) +
  coord_fixed() +
  scale_x_continuous(
    breaks = c(0, 11, 29, 45, 65, 82, 100, 119, 126),
    labels =
      c(
        "Melody left hand",
        "melody 1 ",
        "Melody 2",
        "Melody",
        "Melody 1, octave higher",
        "Melody 2, octave higher",
        "Melody 3, octave higher",
        "fades out",
        ""
      ),
  ) +
  scale_y_continuous(
    breaks = c(0, 11, 31, 49, 69, 90, 110, 131, 140),
    labels =
      c(
        "Melody left hand",
        "melody 1 ",
        "Melody 2",
        "Melody",
        "Melody 1, octave higher",
        "Melody 2, octave higher",
        "Melody 3, octave higher",
        "fades out",
        ""
      ),
  ) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(x = "Lavinia meijers", y = "Yann Tierssen")
maria
```

***

This is a visualization of two performances of the piano piece " Comptine d'un autre été - l'après midi" a performance of Lavina meijers, and the original performance by Yann Tierssen. I chose this music piece because the version of Lavina Meijers was in my playlist, and the version by Yann Tierssen was in the Classical study music playlist. 

It's the soundtrack of the French film Amelie and became popular ever since. It's a piano piece and notice that the left melody of the left hand part never changes (except from intensity). The right hand part has three distinctive melodies in a row that have different characteristics: The first one is catchy and has some mild dissonance, the second has fewer notes and leaves more space, and the third is a flurry of repeated triplets. Halfway through the song, the right hand jumps up an octave and plays the exact same thing as it did in the first half.

The version of Yann Tierssen is 16 seconds longer than the version of lavina meijers (2:22, 2:06). I have to look at his again but I can't distinguish this immediately. I think this is not that that the version of Lavina is a bit faster but its more that it starts earlier and has last pause between the different pieces of the music/less fade. This explains the repetition that we can clearly see in the plot. 

### The Godfather finale [chroma and timbre features]

```{r godfather}
bzt <-
  get_tidy_audio_analysis("78Q2Em989i9lagCBf8XGyX") %>% 
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "mean"
      )
  )
bind_rows(
  bzt %>% 
    compmus_self_similarity(pitches, "cosine") %>% 
    mutate(d = d / max(d), type = "Chroma"),
  bzt %>% 
    compmus_self_similarity(timbre, "euclidean") %>% 
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>% 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_classic() + 
  labs(x = "", y = "")
```

***

The two self-similarity matrices at the right, each summarized at the bar level but with axes in seconds, illustrate pitch- and timbre-based self-similarity within Nini Rotta’s  famous The Godfather Finale. The chroma-based matrix shows us that the piece exists of two parts, divided around the 110 sec. But furthermore the chroma similarities seems to be all over the place, although you can distinguish the violin part (110-160 sec). This part is also very visible in the Timbre-based matrix.
