---
title: "Computational Musicology"
author: "Ragnar van Dort"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: journal
    css: styles.css
    self_contained: false

---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Modules to import
library(ggpubr)
library(cowplot)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(spotifyr)
library(plotly)
library(knitr)
library(DT)
library(showtext)
library(compmus)
library(gridExtra)
library(grid)
# Custom font
font_add("TTNorms", "TTNorms-ExtraBold.otf")
showtext_auto()
```

```{r, include=FALSE}
# Load corpus playlist from Spotify API
corpus <- get_playlist_audio_features("", "69WQ6EmXwnePRjBmS6K7ic") %>%
  mutate(mode = ifelse(mode == 0, "Minor", "Major")) %>% 
  mutate(artists.name = map_chr(track.album.artists, function(x) x$name[1]))
```

```{r, include=FALSE}
# Define your own theme function below
theme_vapor <- function() {
    theme_minimal() +
    theme(
      text = element_text(family = "TTNorms", color = "#e7e247", size=14),
      plot.title = element_text(hjust = 0.5),
      plot.background = element_rect(fill = "#181818", colour="#181818"),
      panel.background  = element_rect(fill = "#989898"),
      legend.key = element_rect(fill = "#181818"),
      #legend.background = element_rect(fill = "#131313")
      panel.grid = element_line(color = "#e7e247")
    )
}
```

### Comparing GoldBand vs De jeugd van tegenwoordig 

**Is Goldband de nieuwe jeugd van tegenwoordig?** 

Goldband wordt soms de nieuwe jeugd van tegenwoordig genoemd, en in veel opzichten lijken ze ook op elkaar

```{=html}
<iframe style="border-radius:12px" src="https://open.spotify.com/embed/playlist/05LvyQi0PJYjmzdAMJn3oy?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>
```
<iframe style="border-radius:12px" src="https://open.spotify.com/embed/playlist/2BzEBqnH9kKBVedAeRAG3h?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>

### Track level features {data-commentary-width=400}

#### __Energy and Danceability __

```{r, echo = FALSE}
# color = mode
p <- ggplot(corpus, aes(energy, danceability, label = track.name, color = artists.name)) + geom_point(alpha=.6, size=3) + scale_fill_discrete() +
  labs(
  x="Energy",
  y="Danceability",
  title="Danceability & Energy",
  caption="Source: Spotify API",
  color = "Artist(s)                       ") +
  xlim(0,1) + 
  ylim(0,1) +
  theme_vapor()
myplot = ggplotly(p)
# fixing weird legend formatting: https://stackoverflow.com/questions/49133395/strange-formatting-of-legend-in-ggplotly-in-r
for (i in 1:length(myplot$x$data)){
    if (!is.null(myplot$x$data[[i]]$name)){
        myplot$x$data[[i]]$name =  gsub("\\(","",str_split(myplot$x$data[[i]]$name,",")[[1]][1])
    }
}
myplot
```

**What is the vibe?**

Energy and Danceability can convey the mood of songs. We can see clearly that that both artists score pretty high on the features in general

#### __The valence in the different playlistst__

```{r some plots, echo=FALSE}
goldband <- get_playlist_audio_features("", "05LvyQi0PJYjmzdAMJn3oy")
jeugd <- get_playlist_audio_features("", "2BzEBqnH9kKBVedAeRAG3h")

study <-
  goldband %>%
  mutate(country = "Goldband") %>%
  bind_rows(jeugd %>% mutate(country = "De jeugd van tegenwoordig")) %>%
  mutate(
    country = fct_relevel(country, "Goldband, De jeugd van Tegenwoordig")
  )
green <- "#1ed760"
pink <- "#eb34b1"
pink <- "#ff6f59"
blue <- "#17bebb"


viz4 <- ggplot(study, aes(x=valence, fill=playlist_name,
                            text = paste(playlist_name)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c(blue, pink))+
  labs(x="Valence", y="Density") +
  guides(fill=guide_legend(title="country"))+
  theme_minimal()+
  ggtitle("Distribution of Valence")+
  theme_vapor()

ggplotly(viz4, tooltip=c("text"))

```

**Valence in the playlists** 
tekst tekst tekst

### Dynamic Time Warping {data-commentary-width=400}

```{r, echo = FALSE}
## Sterrenstof de jeugd
crock_original <-
  get_tidy_audio_analysis("7D5vAulNfrQV6xEwzgH0OF") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
## Sterrenstof guus
crock_remaster <-
  get_tidy_audio_analysis("3JmXdmtkkbjT18dbuPIHRQ") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
dynamic_timewarp <- compmus_long_distance(
  crock_original %>% mutate(pitches = map(pitches, compmus_normalise, "chebyshev")),
  crock_remaster %>% mutate(pitches = map(pitches, compmus_normalise, "chebyshev")),
  feature = pitches,
  method = "euclidean"
) %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_equal() +
  labs(x = "Sterrenstof-De jeugd van tegenwoordig", y = "Sterrenstof-Guus Meeuwis", 
       title = "Dynamic Timewarp") +
  theme_vapor() + 
  scale_fill_viridis_c(guide = NULL)
# Fixing white backdrop: https://stackoverflow.com/questions/57051313/coord-map-of-ggplot2-ignores-plot-background-and-produces-white-margins-how
gt <- ggplotGrob(dynamic_timewarp)
grid.newpage()
grid.draw(rectGrob(gp = gpar(fill = "#181818", col = "#181818")))
grid.draw(gt)
```

***

```{=html}
<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/3TnJ7M6in8Pb5EyGBUK02Y?utm_source=generator&theme=0" width="100%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>
```


<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/5srgpROEtK8KLKQl5FW7Ub?utm_source=generator&theme=0" width="100%" height="80" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>


**Comparing sterrenstofr**

Probably the most popular song of De Jeugd van tegenwoordig is without a doubt Sterrenstof. For there 10 year jubileum...

### Chroma and Tibre features {data-commentary-width=400}

#### __JAJA NEENEE__
```{r, echo = FALSE}
jajaneenee <-
  get_tidy_audio_analysis("6UjBPdD3La0j1JTaT7R6fK") %>% 
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "mean"
      )
  )
bind_rows(
  jajaneenee %>% 
    compmus_self_similarity(pitches, "cosine") %>% 
    mutate(d = d / max(d), type = "Chroma"),
  jajaneenee %>% 
    compmus_self_similarity(timbre, "euclidean") %>% 
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>% 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_vapor() + theme(strip.text = element_text(colour = "#eb34b1"), panel.background  = element_rect(fill = "#121212", colour="#121212")) + 
  labs(x = "", y = "", title = "JaJa NeeNee")
```
**Test**

I wanted to explore the song 'Turbo Killer' by 'Carpenter Brut'. This song really stands out for quick pacing and constantly building and escalating upon the previous 'verse'. The song gives a sense of progression or fast movement/speed, and that is probably why this is one of my favorite tracks in this corpus.

#### __MANON__
```{r, echo = FALSE}
manon <-
  get_tidy_audio_analysis("7FGjTXmgLoQ9kNiD5rSmRs") %>% 
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "mean"
      )
  )
bind_rows(
  manon %>% 
    compmus_self_similarity(pitches, "cosine") %>% 
    mutate(d = d / max(d), type = "Chroma"),
  manon %>% 
    compmus_self_similarity(timbre, "euclidean") %>% 
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>% 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_vapor() + theme(strip.text = element_text(colour = "#eb34b1"), panel.background  = element_rect(fill = "#121212", colour="#121212")) + 
  labs(x = "", y = "", title = "Manon")
```
**The Chroma matrix**

When looking at the Chroma matrix we see a lot of tiny changes in the first forty seconds. Every seven-ish seconds there is a change. After 40 seconds the song changes into high tempo guitar only, and the following blocks all add additional elements to this. The matrix turns out very interesting, because you can see how the song constantly builds up to more complexity.

